---
title: "R Notebook"
output: html_notebook
---

Load libraries and public variables
```{r, include=FALSE} 
# This chunk will be evaluated but the output is suppressed
rmarkdown::render("../../publicVariables/createPublicVariables.Rmd")
```

```{r}
# Raw simulation output folder
rawSimOutFolder <- "/team/batch_SMOTNT/experiment1_output"
simOutputH5 <- paste0("../simOutputH5Data/", experimentID, "_output.h5")
```

### Step 1: Write the groups into the HDF5 file(<1 min to run) for simOutputH5
```{r}
h5createFile(simOutputH5)
fileHandle <- H5Fopen(simOutputH5)

# Create 1st + 2nd level groups for dtype1/2.
# dtype1 and 2 are combined here b/c their 1st/2nd dirs are same. But in the next chunk, when creating datasets, their dimensions are different, so the datasets need to be created and written in separately
i=1
for(gene in geneNames){ 
    # Keepping track of the progress, i max = geneNumbers
    if(i%%1000==0){
        print(i)
    }
  
    # 1st level dir: 4839 genes
    h5createGroup(fileHandle, gene)  
    
    # 2nd level dir: data types out of dtype1 and dtype2
    dtpList <- paste(gene, c(dtype1, dtype2), sep="/")
        for(dtp in dtpList){          
        h5createGroup(fileHandle, dtp)
    }
    i=i+1
}

# Create 1st and 2nd level groups for dtype3
for(dtp in dtype3)
{ 
    if(dtp == "avg_ribo_tRNA"){
        # 1st level dir: dtype3
        h5createGroup(fileHandle, dtp)  
    }
    else{
        # 1st level dir: dtype3
        h5createGroup(fileHandle, dtp)  
        dc.r.List <- paste(dtp, dc.r.df$paraCombo, sep="/")   
          
        # 2nd level dir: dc/r combo or mRNAconstant
        for(dc.r in dc.r.List){ 
            h5createGroup(fileHandle, dc.r)
        }
    }
}



h5closeAll()

# To check the h5 file structures:
####################################### dtype1
# h5read(simOutputH5,'YAL001C/')
# h5read(simOutputH5,'YAL001C/gene_sp_countElng_perMin')
# h5read(simOutputH5,'YAL001C/gene_sp_decCount_perMin')
# h5read(simOutputH5,'YAL001C/gene_sp_meanElngTime_perMin')
# h5read(simOutputH5,'YAL001C/gene_sp_mRNAcount_perMin')
# h5read(simOutputH5,'YAL001C/gene_sp_mRNAmarkedDecay_perMin')
# h5read(simOutputH5,'YAL001C/gene_sp_synCount_perMin')
# h5read(simOutputH5,'YAL001C/gene_sp_totBoundRibo_perMin')
# h5read(simOutputH5,'YAL001C/gene_sp_varianceElngTime_perMin')
# h5read(simOutputH5,'YAL001C/gene_sp_varianceBoundRibo_perMin')

####################################### dtype2
# h5read(simOutputH5,'YAL001C/gene_initimes')
# h5read(simOutputH5,'YAL001C/gene_totetimes')
# h5read(simOutputH5,'YAL001C/mRNA_lifeTimes')

####################################### dtype3
# h5read(simOutputH5,'etimes')
# h5read(simOutputH5,'codon_sp_elngRibo_perMin')
# h5read(simOutputH5,'free_ribo_tRNA_perMin')
# h5read(simOutputH5,'avg_ribo_tRNA')

```

### Step 2: Create empty compressed dataset in h5, total ~1h, 558Mb total;;; 
```{r}
fileHandle <- H5Fopen(simOutputH5)

####################################### create datasets for dtype1
# dtype1: 1st level: 4839 genes, 2nd: dtype, 3rd: dc/r combo or mRNAconstant, 4th:50 rep X 720 min

# 1st: 4839 genes
for(i in 1:geneNumbers){ 
    # keeping track of the progress
    if(i%%500==0){
        print(paste("dtype1", i, Sys.time()))
    } 
  
    # 2nd: data type out of dtype1
    for(dtp in dtype1){
        # 3rd: dc/r combo or mRNAconstant
        for (dc.r in dc.r.df$paraCombo){
            groupname <- paste(geneNames[i], dtp, dc.r, sep="/")
            h5createDataset(file = fileHandle,
                            dataset = groupname,
                            dims = c(numTechReps, simTimeMin), 
                            storage.mode = "double",
                            chunk = c(numTechReps, simTimeMin),
                            level = 6) # intermediate level=6 for the balance between compression size and speed
        }
    }
}

####################################### create datasets for dtype2
# dtype2: 1st/2nd/3rd directory are the same as dtype1
# dtype2 <- c("gene_initimes",          # 4th:50 rep X 3 (#events, mean, var), Average time between initiation of all genes
#             "gene_totetimes",         # 4th:50 rep X 3 (#events, mean, var), Average total elongation times of all genes
#             "mRNA_lifeTimes")         # 4th:50 X 100 (=5000 life time data point, some spots will be filled with NAs since not all genes would have 5000 life times collected). 

# 1st: 4839 genes
for(i in 1:geneNumbers){
    # keeping track of the progress
    if(i%%500==0){
        print(paste("dtype2", i, Sys.time()))
    }
    
    # 2nd: datatype out of dtype2
    for(dtp in dtype2){
        ncoldtype2 <- ifelse(dtp=="mRNA_lifeTimes", 100, 3)
        # 3rd: dc/r combo or mRNAconstant
        for (dc.r in dc.r.df$paraCombo){
            groupname <- paste(geneNames[i], dtp, dc.r, sep="/")  
            h5createDataset(file = fileHandle,
                            dataset = groupname,
                            dims = c(numTechReps, ncoldtype2), 
                            storage.mode = "double",
                            chunk = c(numTechReps, ncoldtype2),
                            level = 6)
        }
    }
}

####################################### create datasets for dtype3
# dtype3: 1st: dtype, 2nd: dc/r combo or mRNAconstant
# dtype3 <- c("etimes",                   # 3rd: 61cid, 4th: 50rep X 2(#events, mean), Elongation times of all codons
#             "codon_sp_elngRibo_perMin", # 3rd: 61cid, 4th: 50rep X 720min, the number of elongatable ribosomes per codon type at each minute mark 
#             "free_ribo_tRNA_perMin",    # 3rd: ribo + 41 tid, 4th: 50rep X 720min, the free ribosome number and free tRNA number per tRNA type at each minute mark
#             "avg_ribo_tRNA")            # 3rd: 50rep X 42(41ribo, 1tRNA), Average number of free ribosomes and tRNAs at equilibrium
             
ncoldtype3 <- c(2, 720, 720, 42)  # ncol for each of dtype3
i=1 # iterate through the four numbers in ncoldtype3
# 1st: dtype
for(dtp in dtype3){ 
    # keep track of the progress
    print(paste("dtype3", dtp, Sys.time()))
  
    # 2nd:dc/r combo or mRNAconstant
    for(dc.r in dc.r.df$paraCombo){
        if(dtp=="etimes" || dtp=="codon_sp_elngRibo_perMin"){
            # j for 61 codon id, corresponding to the cid in [0,60] in the source program
            for(j in 0:60){
                groupname <- paste(dtp, dc.r, paste0("cid", j), sep="/")
                h5createDataset(file = fileHandle,
                                dataset = groupname,
                                dims = c(numTechReps, ncoldtype3[which(dtype3==dtp)]), 
                                storage.mode = "double",
                                chunk = c(numTechReps, ncoldtype3[which(dtype3==dtp)]),
                                level = 6)
            }
        }
        else if(dtp=="free_ribo_tRNA_perMin"){
            groupname <- paste(dtp, dc.r, "ribo", sep="/")
            h5createDataset(file = fileHandle,
                dataset = groupname,
                dims = c(numTechReps, ncoldtype3[which(dtype3==dtp)]), 
                storage.mode = "double",
                chunk = c(numTechReps, ncoldtype3[which(dtype3==dtp)]),
                level = 6)
                
            # j for 41 tRNA id, corresponding to the tid in [0,40] in the source program
            for(j in 0:40){
                groupname <- paste(dtp, dc.r, paste0("tid", j), sep="/")
                h5createDataset(file = fileHandle,
                    dataset = groupname,
                    dims = c(numTechReps, ncoldtype3[which(dtype3==dtp)]), 
                    storage.mode = "double",
                    chunk = c(numTechReps, ncoldtype3[which(dtype3==dtp)]),
                    level = 6)
            }
        }
        else{   #"avg_ribo_tRNA"
            groupname <- paste(dtp, dc.r, sep="/")
            h5createDataset(file = fileHandle,
                            dataset = groupname,
                            dims = c(numTechReps, ncoldtype3[which(dtype3==dtp)]),
                            storage.mode = "double",
                            chunk = c(numTechReps, ncoldtype3[which(dtype3==dtp)]),
                            level = 6)
        }
    }
    i=i+1
}

h5closeAll()
# keep track of the progress
print(Sys.time())

# To check the h5 file contents
####################################### dtype1
# h5read(simOutputH5,'YAL001C/gene_sp_countElng_perMin/mRNAconstant')
# h5read(simOutputH5,'YAL001C/gene_sp_countElng_perMin/dc_0_r_0')
# h5read(simOutputH5,'YAL001C/gene_sp_decCount_perMin/dc_0.2_r_0.1')
# h5read(simOutputH5,'YAL001C/gene_sp_meanElngTime_perMin/dc_0.4_r_0.4')
# h5read(simOutputH5,'YAL001C/gene_sp_mRNAcount_perMindc_0.4_r_1')
# h5read(simOutputH5,'YAL001C/gene_sp_mRNAmarkedDecay_perMin/dc_0.6_r_0.1')
# h5read(simOutputH5,'YAL001C/gene_sp_synCount_perMin/dc_0.6_r_0.4')
# h5read(simOutputH5,'YAL001C/gene_sp_totBoundRibo_perMin/dc_0.8_r_0.4')
# h5read(simOutputH5,'YAL001C/gene_sp_varianceElngTime_perMin/dc_1_r_0')
# h5read(simOutputH5,'YAL001C/gene_sp_varianceBoundRibo_perMin/dc_1_r_0.4')

####################################### dtype2
# h5read(simOutputH5,'YAL001C/gene_initimes/dc_0.4_r_0.1')
# h5read(simOutputH5,'YAL002W/gene_totetimes/dc_0.8_r_1')
# h5read(simOutputH5,'YAL003W/mRNA_lifeTimes/dc_1_r_0.4')

####################################### dtype3
# h5read(simOutputH5,'etimes/mRNAconstant')
# h5read(simOutputH5,'codon_sp_elngRibo_perMin/dc_0_r_0.1')
# h5read(simOutputH5,'free_ribo_tRNA_perMin/dc_0.2_r_1')
# h5read(simOutputH5,'avg_ribo_tRNA/dc_0.6_r_0.4')
```


### Step 3: Prep for writing data for dtype1/2 (took ~1min to run), this step should be ran after all output files are ready stored on the server, the filenames will be stored in each fileGroup.
```{r}
createFileGroups <- function(numberOfDataType, datatype, dcrDataframe){  #numberOfDataType = 9/3/4 for dtype1/2/3
    numGroups <- numberOfDataType * length(dcrDataframe$paraCombo)   
    # e.g. numGroups: for dtype1=9*25=225, dtype2=3*25=75, dtype3=4*25=100
    
    # fileGroupIndex contains the indices of combinations between data type and dc.r.combo. 
    fileGroupIndex <- tibble(data.frame(dataType = integer(numGroups), 
                                 dc.r.combo = integer(numGroups)))
    groupid = 1
    for(i in 1:numberOfDataType){ #-------- 9 data types included in both dtype1 e.g. meanElngTime, etc
        for(j in 1:length(dcrDataframe$paraCombo)){ #-----------------  dc/r combo + 1 mRNAconstant
            if(groupid <=numGroups){
                fileGroupIndex[groupid, ] <- as.list(c(i, j)) 
                # as.list() otherwise gets error message "Row updates require a list value"
                groupid <- groupid+1
            }
        }
    }
    
    # fileGroup contains names of all the raw output files corresponding to each combination of data type X dc.r combination
    fileGroup = tibble(data.frame(matrix(NA, ncol = 2 + numTechReps, nrow = numGroups)))
    for(groupID in 1:numGroups){
        dtp <- as.character(datatype[unlist(fileGroupIndex[groupID, 1])])  #e.g. "gene_sp_countElng_perMin"
        dc.r <- (dcrDataframe$paraCombo)[unlist(fileGroupIndex[groupID, 2])] # e.g. "dc_0_r_1"
        outfolder <- paste(rawSimOutFolder, dtp, sep="/")
        
        if(dc.r=="mRNAconstant"){
            fileGroupVec <- paste0(dtp, "/", list.files(outfolder, pattern = glob2rx(paste0("*", "mRNAconstant", "*", dtp, "*"))))
        }else{
            # fileGroupVec contains 2 individual info + N files that has same data type and dc/r combo but N different seeds. Adding "_allGenesDe" to prevent from including the allGenesCtrl files when dc.r="dc_0.2_r_0.1"
            fileGroupVec <- paste0(dtp, "/", list.files(outfolder, pattern = glob2rx(paste0("*_", dc.r, "_allGenesDe*", dtp, "*")))) 
        }
        
        fileGroupVec <- c(dtp, dc.r, fileGroupVec)
        fileGroup[, ] <- as.list(fileGroupVec)
        
        # keep track of the progress
        if(groupID%%50==0){  
            print(groupID)
        }
    }
    
    
    save(fileGroup, fileGroupIndex, file=paste0("../fileGroup_", deparse(substitute(datatype)), ".rdata"))
    
}

createFileGroups(length(dtype1), dtype1, dc.r.df)
createFileGroups(length(dtype2), dtype2, dc.r.df)
createFileGroups(length(dtype3), dtype3, dc.r.df)

```

### Step 4: Functions to write data from raw output into h5
```{r}
####################################################################### dtype1
# let N output files(for N # of seeds/simulation technical replicates) be a group, then there are a total of length(dtype1) * length(dc.r.df$paraCombo) = 9*25=225 groups for dtype1
# each of these N files (i.e. each groupId) have the same dataType + dc/r combo or mRNAconstant
write_h5_dtype1 <- function(groupID, fileGroup){            # input = groupID 1:225
    dtp <- as.character(unlist(fileGroup[groupID, 1]))      # e.g. "gene_sp_varianceBoundRibo_perMin"
    dc.r <- as.character(unlist(fileGroup[groupID, 2]))     # e.g. "dc_0_r_1"
  
    # Step 1: read all the file contents for these N files into a list
    currentNlist <- list()  
    for(ii in 1:numTechReps){
        # large size list. e.g. 159Mb for 10 seeds
        # dim(currentNlist[[1]]) = 720 4840, 4840 = time mark + 4839 genes
        # read_delim always returns an extra column???
        currentNlist[[ii]] = read.table(file=paste0(rawSimOutFolder, "/", as.character(unlist(fileGroup[groupID, ii+2])))) 
    }
    
    # Step 2: read from each col of each file (one element of the list) into each row of dataNseedsH5
    for(i in 2:(geneNumbers+1)){  # 2:4840, 1st col=simulation time in seconds, rest col = 4839 genes, total col=time mark + 4839 genes = 4840
        # dataNseedsH5 contains the data of all N seeds for a given gene(1 of 4839)
        dataNseedsH5 <- matrix(NA, nrow = numTechReps, ncol = simTimeMin)
        
        for(j in 1:numTechReps){  # indexes for data of n seeds from n files
          dataNseedsH5[j, ] <- currentNlist[[j]][, i]
        }
        
        groupname <- paste(geneNames[i-1], dtp, dc.r, sep="/")
        h5write(dataNseedsH5, fileHandle, name = groupname, start=c(1, 1))
      
    }
    print(paste("dtype1", groupID, Sys.time()))
}  

####################################################################### dtype2
write_h5_dtype2 <- function(groupID, fileGroup){            # input = groupID 1:75, 3*25=75
    dtp <- as.character(unlist(fileGroup[groupID, 1]))
    dc.r <- as.character(unlist(fileGroup[groupID, 2]))
    
    if(dtp != "mRNA_lifeTimes"){
        # Step 1: read all the file contents for these N files into a list
        currentNlist <- list()  
        for(ii in 1:numTechReps){
            # large size list. e.g. 1.1Mb
            # dim(currentNlist[[1]]) = 4840 4or6 
            currentNlist[[ii]] <- read.table(file=paste0(rawSimOutFolder, "/", as.character(unlist(fileGroup[groupID, ii+2]))),  header=T)
        }
      
        # Step 2: read from each col of each file(one element of the list) into each row of dataNseedsH5
        # raw output files in dtp2 (except for mRNA_lifeTimes) all have 4 columns: geneNumber, numberOfEvents, mean, variance. Below is to set the start and end col as 2 and 4, excluding 1st the geneNumber column.
        startcol_dtp2 <- 2
        endcol_dtp2 <- 4
        ncol_dtp2 <- 3 
        for(i in 1:geneNumbers){  # indexes for all genes, each row is a gene in raw output file
            # dataNseedsH5 contains the data of all N seeds for a given gene(1 of 4839)
            dataNseedsH5 <- matrix(NA, nrow = numTechReps, ncol = ncol_dtp2)

            for(j in 1:numTechReps){  # indexes for data of n seeds from n files
                dataNseedsH5[j, ] <- as.numeric(currentNlist[[j]][i, startcol_dtp2:endcol_dtp2])
            }
            
            groupname <- paste(geneNames[i], dtp, dc.r, sep="/")
            h5write(dataNseedsH5, fileHandle, name = groupname, start=c(1, 1))
            
        }
    }
    
    else if(dtp=="mRNA_lifeTimes"){
        # Step 1: read all the file contents for these N files into a list
        currentNlist <- list()  
        for(ii in 1:numTechReps){
            # large size list. e.g. 20Mb*100~2Gb
            # using readLines() instead of read.table() because each line of the raw output contains different number of items (because different genes has recorded different number of lifetimes)
            currentNlist[[ii]] <- readLines(con=paste0(rawSimOutFolder, "/", as.character(unlist(fileGroup[groupID, ii+2]))))  
        }
        
        # Step 2: read from each line of each file(one element of the list), sample 5,000 numbers for each gene, then write into dataNseedsH5
        for(i in 1:geneNumbers){  # indexes for all genes, each row is a gene in original file
            lifeTimeVec = c()
            for(j in 1:numTechReps){  # indexes for data of n seeds from n files
                 fcontent <- as.numeric(unlist(strsplit(currentNlist[[j]][i], " ")))
                 # in raw output file, 1st number=gid, 2nd number= number of lifeTimes, from 3rd on are the life times.
                 lifeTimeVec <- c(lifeTimeVec, fcontent[3:(fcontent[2]+2)])
            }
            
            # If the lifeTimeVec length is smaller than numTechReps*100=50*100=5000, then supplement sampledLifeTimes with multiple NAs so it can fill up dataNseedsH5.
            if(length(lifeTimeVec)>numTechReps*100){
                sampledLifeTimes <- sample(lifeTimeVec, numTechReps*100)
            }else{
                sampledLifeTimes <- c(lifeTimeVec, rep(NA, numTechReps*100-length(lifeTimeVec)))
            }
          
            # dataNseedsH5 contains the data of all N seeds for a given gene(1 of 4839)
            dataNseedsH5 <- matrix(sampledLifeTimes, nrow = numTechReps,  ncol = 100)
            groupname <- paste(geneNames[i], dtp, dc.r, sep="/")
            h5write(dataNseedsH5, fileHandle, name = groupname, start=c(1, 1))
        }
    }
    print(paste("dtype2", groupID, Sys.time()))
}  

####################################################################### dtype3
write_h5_dtype3 <- function(groupID, fileGroup){  # input = groupID 1:100, 4*25=100
    ncoldtype3 <- c(2, 720, 720, 42)  # ncol for each of dtype3. Refer to "../../publicVariables/createPublicVariables.Rmd" for the specifications
    dtp <- as.character(unlist(fileGroup[groupID, 1]))
    dc.r <- as.character(unlist(fileGroup[groupID, 2]))
    
    # Step 1: read all the file contents for these N files into a list
    currentNlist <- list()  
    for(ii in 1:numTechReps){
        # large size list. e.g. 1.1Mb
        # dim(currentNlist[[1]]) = 4840 4or6 
        currentNlist[[ii]] <- read.table(file=paste0(rawSimOutFolder, "/", as.character(unlist(fileGroup[groupID, ii+2]))), header=ifelse(dtp=="etimes", TRUE, FALSE))   
    }
  
    # Step 2: read from each col of each file(one element of the list) into each row of dataNseedsH5
    if(dtp=="avg_ribo_tRNA"){
        # dataNseedsH5 contains the data of all N seeds
        dataNseedsH5 <- matrix(NA, nrow = numTechReps, ncol = ncoldtype3[which(dtype3==dtp)])
        for(j in 1:numTechReps){  # indexes for data of n seeds from n files
            dataNseedsH5[j, ] <- as.numeric(currentNlist[[j]][, 2])
        }
        groupname <- paste(dtp, dc.r, sep="/")
        h5write(dataNseedsH5, fileHandle, name = groupname, start=c(1, 1))
    }
    
    else if(dtp=="codon_sp_elngRibo_perMin"){
        for(i in 0:60){
            # dataNseedsH5 contains the data of all N seeds
            dataNseedsH5 <- matrix(NA, nrow = numTechReps, ncol = ncoldtype3[which(dtype3==dtp)])
            for(j in 1:numTechReps){  # indices for data of n seeds from n files
                dataNseedsH5[j, ] <- as.numeric(currentNlist[[j]][, i+2])
            }
            
            groupname <- paste(dtp, dc.r, paste0("cid", i), sep="/")
            h5write(dataNseedsH5, fileHandle, name = groupname, start=c(1, 1))
        }
    }
    
    else if(dtp=="etimes"){
        for(i in 0:60){
            # dataNseedsH5 contains the data of all N seeds
            dataNseedsH5 <- matrix(NA, nrow = numTechReps, ncol = ncoldtype3[which(dtype3==dtp)])
            for(j in 1:numTechReps){  # indexes for data of n seeds from n files
                dataNseedsH5[j, ] <- as.numeric(currentNlist[[j]][i+1, 2:3])
            }
            
            groupname <- paste(dtp, dc.r, paste0("cid", i), sep="/")
            h5write(dataNseedsH5, fileHandle, name = groupname, start=c(1,1))
        }
    }

    else{   #"free_ribo_tRNA_perMin"
        # dataNseedsH5 contains the data of all N seeds
        dataNseedsH5 <- matrix(NA, nrow = numTechReps, ncol = ncoldtype3[which(dtype3==dtp)])
        for(j in 1:numTechReps){  # indexes for data of n seeds from n files
            dataNseedsH5[j, ] <- as.numeric(currentNlist[[j]][, 2])
        }
        
        groupname <- paste(dtp, dc.r, "ribo", sep="/")
        h5write(dataNseedsH5, fileHandle, name = groupname, start=c(1, 1))
        
        for(i in 0:40){
            # dataNseedsH5 contains the data of all N seeds
            dataNseedsH5 <- matrix(NA, nrow = numTechReps, ncol = ncoldtype3[which(dtype3==dtp)])
            for(j in 1:numTechReps){  # indexes for data of n seeds from n files
                dataNseedsH5[j, ] <- as.numeric(currentNlist[[j]][, i+3])
            }
            
            groupname <- paste(dtp, dc.r, paste0("tid", i), sep="/")
            h5write(dataNseedsH5, fileHandle, name = groupname, start=c(1, 1))
        }
    }
    
    print(paste("dtype3", groupID, Sys.time()))
}  

```



### Step 5: Write data from raw output into h5 for dtype1
9:22pm begin, 5:40am ended next morning. took ~8.5h, h5 starts in 558Mb, and ended up in 82G
```{r}
# load environmental variables: fileGroup, fileGroupIndex
load("../fileGroup_dtype1.rdata") 

fileHandle <- H5Fopen(simOutputH5)
for (i in 1:225){   #1:225
    write_h5_dtype1(i, fileGroup = fileGroup)
}

h5closeAll()

####################################### dtype1
# View(h5read(simOutputH5,'YAL001C/gene_sp_countElng_perMin/mRNAconstant'))
# View(h5read(simOutputH5,'YAL001C/gene_sp_countElng_perMin/dc_0_r_0'))
# View(h5read(simOutputH5,'YAL001C/gene_sp_decCount_perMin/dc_0.2_r_0.1'))
# View(h5read(simOutputH5,'YAL001C/gene_sp_meanElngTime_perMin/dc_0.4_r_0.4'))
# View(h5read(simOutputH5,'YAL001C/gene_sp_mRNAcount_perMindc_0.4_r_1'))
# View(h5read(simOutputH5,'YAL001C/gene_sp_mRNAmarkedDecay_perMin/dc_0.6_r_0.1'))
# View(h5read(simOutputH5,'YAL001C/gene_sp_synCount_perMin/dc_0.6_r_0.4'))
# View(h5read(simOutputH5,'YAL001C/gene_sp_totBoundRibo_perMin/dc_0.8_r_0.4'))
# View(h5read(simOutputH5,'YAL001C/gene_sp_varianceElngTime_perMin/dc_1_r_0'))
# View(h5read(simOutputH5,'YAL001C/gene_sp_varianceBoundRibo_perMin/dc_1_r_0.4'))
```



### Step 6: Write data from raw output into h5 for dtype2
took 2 hours, finished as being 86G
```{r}
# load environmental variables: fileGroup, fileGroupIndex
load("../fileGroup_dtype2.rdata")

fileHandle <- H5Fopen(simOutputH5)
for (i in 1:75){   #1:75
    write_h5_dtype2(i, fileGroup = fileGroup)
}

h5closeAll()

####################################### dtype2
# View(h5read(simOutputH5,'YAL001C/gene_initimes/dc_0.4_r_0.1'))
# View(h5read(simOutputH5,'YAL002W/gene_totetimes/dc_0.8_r_1'))
# View(h5read(simOutputH5,'YAL003W/mRNA_lifeTimes/dc_1_r_0.4'))
```



### Step 7: Write data from raw output into h5 for dtype3
took 3 min, finished as being 86G
```{r}
# load environmental variables: fileGroup, fileGroupIndex
load("../fileGroup_dtype3.rdata")

fileHandle <- H5Fopen(simOutputH5)
for (i in 1:100){  # 1:100
  write_h5_dtype3(i, fileGroup = fileGroup)
}

h5closeAll()

####################################### dtype3
# View(h5read(simOutputH5,'etimes/mRNAconstant'))
# View(h5read(simOutputH5,'codon_sp_elngRibo_perMin/dc_0_r_0.1'))
# View(h5read(simOutputH5,'free_ribo_tRNA_perMin/dc_0.2_r_1'))
# View(h5read(simOutputH5,'avg_ribo_tRNA/dc_0.6_r_0.4'))
```