---
title: "R Notebook"
output: html_notebook
---


Load libraries and public variables
```{r, include = FALSE} 
# This chunk will be evaluated but the output is suppressed
rmarkdown::render("../../publicVariables/createPublicVariables.Rmd")
```

The total number of jobs for experiment1 is 1250
6dc X 4r X 50 reps + 50 mRNA constant = 1200 +50 = 1250 simulations
```{r}
set.seed(0)  
# Seeds are required to feed each simulation. C program can set up to 32767 seeds.
seedvec <- sample(1:32767, 1250, replace = F)  

# Max time to run the simulations are (12h*3600s = 43200s)+1000s burn in = 44200s. for most mRNAs their half lives are <35min =>simu time= 10/(log(2)/(35*60)) = 30296s, setting it to 44200s should therefore sufficiently allow the mRNAs to have enough turnover.
maxTime <- 44200 

# Repeating the dc.r.df the number of times = tech reps
dc.r.rep <- tibble(dc = rep(dc.r.df$dc[2:25], numTechReps),  # 2:25 in order to exclude mRNAconstant
                   r = rep(dc.r.df$r[2:25], numTechReps), 
                   paraCombo = rep(dc.r.df$paraCombo[2:25], numTechReps))
```


Write first line into the job file.
The job file should eventually contain 2 header lines + 3 line per job * 1250 jobs = 3752 lines
```{r}
# File containing Amarel jobs
jobFileAmr <- paste0("amarel_", experimentID, "_jobs.mf")  
# To empty the job file from old content, otherwise contents will add up
system(paste0("cp /dev/null ", jobFileAmr))  
# Lines for setting parameters on Amarel
options <- paste0(" -N 1 -n 1 -c 1 --mem=7G -t 25:00:00 -o stdout/%A.out -e error/%A.err -J ", experimentID, "-")
firstLine <- paste0("BATCH_OPTIONS = -p p_shahlab_1, genetics_1, main", options, "amr\n") 

write(firstLine, file = jobFileAmr, append = T)
```

Writing 6dc x 4r x 50 reps = 1200 mRNAvarying jobs into the job file
## (NOTE: THE RELATIVE PATHS BELOW ARE INCORRECT DURING THIS EDIT FOR EXPERIMENT1, NEEDS TO BE CORRECTED. Dec 1st, 2021)
```{r}
for(i in 1:1200){
    outFileName = paste0("amr.",experimentID, ".seed.", seedvec[i], "_", dc.r.rep$paraCombo[i], "_allGenesDecrEqualsSynrScaling")
    write(paste0( ": ../../publicInput/S.cer.genom ../../publicInput/S.cer.tRNA ../input/allGenesDecrEqualsSynrWithScaling_", dc.r.rep$paraCombo[i], "_S.cer.mRNA.ini.abndc.syn.dec ../ac_amarelCode/wrapper.sh ../source/SMOTNT\n", 
                  "    ./../ac_amarelCode/wrapper.sh -Tt ", maxTime, " -Tb 1000 -R 200000 -t 3300000 -N 4839 -dc ", dc.r.rep$dc[i], " -r ", dc.r.rep$r[i], " -F ../../publicInput/S.cer.genom -C ../../publicInput/S.cer.tRNA ", 
                  "-D ../input/allGenesDecrEqualsSynrWithScaling_", dc.r.rep$paraCombo[i], "_S.cer.mRNA.ini.abndc.syn.dec -s ", seedvec[i], 
                  " -O ../output/", outFileName, 
                  " -p1 -p2 -p3 -p4 -p6 -p7\n"), file = jobFileAmr, append = T)
}
```

Writing 50 mRNAconstant jobs into the job file
```{r}
for(i in 1201:1250){
    outFileName = paste0("amr.", experimentID,".seed.",seedvec[i],"_dc_0_r_0_mRNAconstant")
    write(paste0( ": ../../publicInput/S.cer.genom ../../publicInput/S.cer.tRNA ../input/allGenes_mRNAconstant_S.cer.mRNA.ini.abndc.syn.dec ../ac_amarelCode/wrapper.sh ../source/SMOTNT\n",
                  "    ./../ac_amarelCode/wrapper.sh -Tt ",maxTime," -Tb 1000 -R 200000 -t 3300000 -N 4839 -dc 0 -r 0 -F ../../publicInput/S.cer.genom -C ../../publicInput/S.cer.tRNA ",
                  "-D ../input/allGenes_mRNAconstant_S.cer.mRNA.ini.abndc.syn.dec -s ",seedvec[i],
                  " -O ../output/",outFileName,
                  " -p1 -p2 -p3 -p4 -p6 -p7\n"),file = jobFileAmr,append = T)
}
```

