---
title: "R Notebook"
output: html_notebook
---

Load libraries and public variables
```{r, include = FALSE} 
# This chunk will be evaluated but the output is suppressed
rmarkdown::render("../../publicVariables/createPublicVariables.Rmd")
```


```{r}
# File that contains simulation output data in a HDF5 directory structure specified in "/publicVariables/createPublicVariables.Rmd"
simOutputH5 <- paste0("../../dataProcessing/simOutputH5Data/", experimentID, "_output.h5")
#allSimInputOutput <- read_tsv("../allSimInputOutput.feather")

# This folder contains raw output files from the simulations. This is needed for calculate_para_sp_mcpsr()
rawSimOutFolder <- "/team/batch_SMOTNT/experiment1_output"
```



### Refer to Supplementary Methods for the mathematical equations of each of the following calculated metrics.
```{r}
# mpsr = mean protein synthesis rates;  
calculate_gene_sp_mpsr <- function(geneID, simOutputH5){   
    # outvec has 25 elements. Each element is the mpsr (# of ribos that hop off per min for a gene) for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes
    currentGene <- geneNames[geneID]

    # matrix that contains countElng info for current selected gene in mRNAconstant data
    # ps = protein synthesis
    # mpsr = mean protein synthesis rate
    ps_mRNAconstant <- h5read(simOutputH5, paste0(currentGene, "/gene_sp_countElng_perMin/mRNAconstant"))
    mpsr_mRNAconstant <- mean(ps_mRNAconstant)
    outvec <- c(outvec, mpsr_mRNAconstant)
    
    # loop through the dc/r parameter space to calculate the mean_countElng_mRNAvarying
    for(i in 2:nrow(dc.r.df)){
        ps_mRNAvarying <- h5read(simOutputH5, paste(currentGene, "gene_sp_countElng_perMin", dc.r.df$paraCombo[i], sep = "/"))
        mpsr_mRNAvarying <- mean(ps_mRNAvarying)
        outvec <- c(outvec, mpsr_mRNAvarying)
    }
    
    h5closeAll()
    return(outvec)
}
```


```{r}
# mvps = mean of the (variance in protein synthesis among simu times) among technical replicates 
calculate_gene_sp_mvps <- function(geneID, simOutputH5){  
    # outvec has 25 elements. Each element is the mvps for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes
    currentGene <- geneNames[geneID]
    
    # matrix that contains countElng info for current selected gene in mRNAconstant data
    ps_mRNAconstant <- h5read(simOutputH5, paste0(currentGene, "/gene_sp_countElng_perMin/mRNAconstant"))
    # get variance for each row(rep) and then get the mean for all reps
    mvps_mRNAconstant <- mean(apply(ps_mRNAconstant, 1, var))  
    outvec <- c(outvec, mvps_mRNAconstant)
    
    # loop through the dc/r parameter space 
    for(i in 2:nrow(dc.r.df)){ 
        # ps_mRNAvarying is a 50 X 720 matrix
        ps_mRNAvarying <- h5read(simOutputH5, paste(currentGene, "gene_sp_countElng_perMin", dc.r.df$paraCombo[i], sep = "/"))
        # get variance for each row (rech rep) and then get the mean for all technial reps
        mvps_mRNAvarying <- mean(apply(ps_mRNAvarying, 1, var, na.rm = TRUE))  
        outvec <- c(outvec, mvps_mRNAvarying)
    }
    
    h5closeAll()
    return(outvec)
}
```

```{r}
# cvps = coefficient variation of total protein produced per minute
calculate_gene_sp_cvps <- function(geneID, simOutputH5){  
    # outvec is a vector that has 25 elements. Each element is the mvps for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes
    currentGene <- geneNames[geneID]
    
    # loop through the dc/r parameter space to calculate the mean_countElng_mRNAvarying
    for(i in 1:nrow(dc.r.df)){ 
        ps <- h5read(simOutputH5, paste(currentGene, "gene_sp_countElng_perMin", dc.r.df$paraCombo[i], sep = "/"))
        mpsr <- mean(ps) # mean
        # get standard deviation for each row(rep) and then get the mean for all reps
        sdps <- mean(apply(ps, 1, sd, na.rm = TRUE))  
        # cvps = coefficient of variation of protein synthesis
        cvps <- sdps/mpsr 
        outvec <- c(outvec, cvps)
    }
    
    h5closeAll()
    return(outvec)
}
```


```{r}
# cpsr = cellular protein synthesis rate, this function returns the cpsr for 50 technical replicate that is of the same paraCombo, it is therefore not joined into allSimInputOutput
# pid = parameter id, 1:25, 1 = mRNAconstant, 2:25 = different dc/r combos, output is a number
calculate_para_sp_cpsr <- function(pid){  
    # there are 25 para combo * 50 reps = 1250 files in this folder                              
    countElngFolder <- paste0(rawSimOutFolder, "/gene_sp_countElng_perMin/") 
    
    # N = # of tech reps, here NcountElngFiles contains 50 file contents
    if(pid == 1){
        NcountElngFiles <- list.files(countElngFolder, pattern = glob2rx(paste0("*dc_0_r_0_mRNAconstant*")))
    }else{
        NcountElngFiles <- list.files(countElngFolder, pattern = glob2rx(paste0("*", dc.r.df$paraCombo[pid], "_allGenesDecrEqualsSynrScaling*")))
    }
    
    cpsr.all.techRep <- c()
    for(ii in 1:numTechReps){ 
        # cps = cellular protein synthesis, is a vector that contains 719 numbers(720-1)
        cps = apply(read.table(file = paste0(countElngFolder, NcountElngFiles[ii]))[, 2:4840], 1, sum) # get the sum of each line in a single file
      
        # cps give a vector of 719 numbers, then you get the mean of those 719 numbers
        # cpsr = cellular protein synthesis rate
        cpsr <- mean(cps)
        cpsr.all.techRep <- c(cpsr.all.techRep, cpsr)  # this will contain 50 numbers
    }

    return(cpsr.all.techRep)
}
```


```{r}
# mcpsr = mean cellular protein synthesis rate among technical reps
# parameter id, 1:25, 1 = mRNAconstant, 2:25 = different dc/r combos, output is a number
calculate_para_sp_mcpsr <- function(pid){  
    # there are 25 para combo * 50 reps = 1250 files in this folder                  
    countElngFolder <- paste0(rawSimOutFolder, "/gene_sp_countElng_perMin/") 
    
    # N = # of tech reps, here NcountElngFiles contains 50 file contents
    if(pid == 1){
        NcountElngFiles <- list.files(countElngFolder, pattern = glob2rx(paste0("*dc_0_r_0_mRNAconstant*")))
    }else{
        NcountElngFiles <- list.files(countElngFolder, pattern = glob2rx(paste0("*", dc.r.df$paraCombo[pid], "_allGenesDecrEqualsSynrScaling*")))
    }
    
    cpsr.all.techRep <- c()
    for(ii in 1:numTechReps){
        # cps = cellular protein synthesis, is a vector that contains 720 numbers
        # get the sum of each line in a single file
        cps = apply(read.table(file = paste0(countElngFolder, NcountElngFiles[ii]))[, 2:4840], 1, sum) 
        
        # cpsr = cellular protein synthesis rate = get the mean of those 720 numbers in cps
        cpsr <- mean(cps)
        # this will contain 50 numbers, 1 for each tech rep
        cpsr.all.techRep <- c(cpsr.all.techRep, cpsr)  
    }
    # mean of the cpsr among 50 technical reps
    mcpsr <- mean(cpsr.all.techRep)
    return(mcpsr)
}
```

```{r}
# mte = mean translation efficiency
calculate_gene_sp_mte <- function(geneID, simOutputH5){   
    # outvec has 25 elements. Each element is the mte for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes, and its initial mRNA level
    currentGene <- geneNames[geneID]
    mRNAabundanceCurrentGene <- simInputFeatures$mRNAabundance[geneID]
    
    # matrix that contains countElng info for current gene.
    # te = translation efficiency, mte = mean translation efficiency
    countElng_mRNAconstant <- h5read(simOutputH5, paste0(currentGene, "/gene_sp_countElng_perMin/mRNAconstant"))
    te_mRNAconstant <- countElng_mRNAconstant/mRNAabundanceCurrentGene
    mte_mRNAconstant <- mean(te_mRNAconstant)
    outvec <- c(outvec, mte_mRNAconstant)
    
    # loop through the dc/r parameter space
    for(i in 2:nrow(dc.r.df)){ 
        countElng_mRNAvarying <- h5read(simOutputH5, paste(currentGene, "gene_sp_countElng_perMin", dc.r.df$paraCombo[i], sep = "/"))
        mRNAcount_mRNAvarying <- h5read(simOutputH5, paste(currentGene, "gene_sp_mRNAcount_perMin", dc.r.df$paraCombo[i], sep = "/"))
        # te = translation efficiency
        te_mRNAvarying <- countElng_mRNAvarying/mRNAcount_mRNAvarying   
        # change all the "Inf"s to "Nan"s
        te_mRNAvarying[!is.finite(te_mRNAvarying)] <- NaN 
        # mte = mean translation rate averaged over time and tech reps
        mte_mRNAvarying <- mean(te_mRNAvarying, na.rm = TRUE) 
        outvec <- c(outvec, mte_mRNAvarying)
    }
    
    h5closeAll()
    return(outvec)
}
```

```{r}
# mvte = mean of the variance in translation efficiency
calculate_gene_sp_mvte <- function(geneID, simOutputH5){     
    # outvec has 25 elements. Each element is the mvte for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes, and its initial mRNA level
    currentGene <- geneNames[geneID]
    mRNAabundanceCurrentGene <- simInputFeatures$mRNAabundance[geneID]
    
    # matrix that contains countElng info for current selected gene
    countElng_mRNAconstant <- h5read(simOutputH5, paste0(currentGene, "/gene_sp_countElng_perMin/mRNAconstant"))
    te_mRNAconstant <- countElng_mRNAconstant/mRNAabundanceCurrentGene
    # get variance for each row(rep) and then get the mean for all reps
    mvte_mRNAconstant <- mean(apply(te_mRNAconstant, 1, var, na.rm = TRUE))  
    outvec <- c(outvec, mvte_mRNAconstant)
    
    # loop through the dc/r parameter space
    for(i in 2:nrow(dc.r.df)){  
        countElng_mRNAvarying <- h5read(simOutputH5, paste(currentGene, "gene_sp_countElng_perMin", dc.r.df$paraCombo[i], sep = "/"))
        mRNAcount_mRNAvarying <- h5read(simOutputH5, paste(currentGene, "gene_sp_mRNAcount_perMin", dc.r.df$paraCombo[i], sep = "/"))
        # te = translation rate
        te_mRNAvarying <- countElng_mRNAvarying/mRNAcount_mRNAvarying   
        # change all the "Inf"s to "Nan"s
        te_mRNAvarying[!is.finite(te_mRNAvarying)] <- NaN 
        # get var for each row(rep) and then get the mean for all reps
        mvte_mRNAvarying <- mean(apply(te_mRNAvarying, 1, var, na.rm = TRUE))  
        outvec <- c(outvec, mvte_mRNAvarying)
    }
    
    h5closeAll()
    return(outvec)
}
```




```{r}
# mmc = mean mRNA count
calculate_gene_sp_mmc <- function(geneID, simOutputH5){   
    # outvec has 25 elements. Each element is the mmc for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes,  and its initial mRNA level
    currentGene <- geneNames[geneID]
    mRNAabundanceCurrentGene <- simInputFeatures$mRNAabundance[geneID]
    outvec <- c(outvec, mRNAabundanceCurrentGene)
    
    # loop through the dc/r parameter space
    for(i in 2:nrow(dc.r.df)){  
        mRNAcount_mRNAvarying <- h5read(simOutputH5, paste(currentGene, "gene_sp_mRNAcount_perMin", dc.r.df$paraCombo[i], sep = "/"))
        mmc_mRNAvarying <- mean(mRNAcount_mRNAvarying)
        outvec <- c(outvec, mmc_mRNAvarying)
    }
    
    h5closeAll()
    return(outvec)
}
```

```{r}
# mvmc = mean of the variance in mRNA count
calculate_gene_sp_mvmc <- function(geneID, simOutputH5){    
    # outvec has 25 elements. Each element is the mvmc for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes
    currentGene <- geneNames[geneID]

    mvmc_mRNAconstant <- 0  # the variance in mRNA count for mRNAconstant is 0
    outvec <- c(outvec, mvmc_mRNAconstant)
    
    # loop through the dc/r parameter space, excluding mRNAconstant
    for(i in 2:nrow(dc.r.df)){  
        mc_mRNAvarying <- h5read(simOutputH5, paste(currentGene, "gene_sp_mRNAcount_perMin", dc.r.df$paraCombo[i], sep = "/"))
        # get variance for each row(rep) and then get the mean for all reps
        mvmc_mRNAvarying <- mean(apply(mc_mRNAvarying, 1, var, na.rm = TRUE))  
        outvec <- c(outvec, mvmc_mRNAvarying)
    }
    
    h5closeAll()
    return(outvec)
}
```

```{r}
# cvmc = coefficient variation of mRNA count
calculate_gene_sp_cvmc <- function(geneID, simOutputH5){    
    # outvec has 25 elements. Each element is the cvmc for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes, and its initial mRNA level
    currentGene <- geneNames[geneID]
    # 0 to represent mRNAconstant
    outvec <- c(outvec, 0)
    
    # loop through the dc/r parameter space, excluding mRNAconstant
    for(i in 2:nrow(dc.r.df)){  
        mRNAcount_mRNAvarying <- h5read(simOutputH5, paste(currentGene, "gene_sp_mRNAcount_perMin", dc.r.df$paraCombo[i], sep = "/"))
        # mmc = mean mRNA count
        mmc_mRNAvarying <- mean(mRNAcount_mRNAvarying) 
        # sdmc = standard deviation for each row(rep) and then get the mean for all reps
        sdmc_mRNAvarying <- mean(apply(mRNAcount_mRNAvarying, 1, sd, na.rm = TRUE)) 
        # cvmc = coefficient of variation
        cvmc_mRNAvarying <- sdmc_mRNAvarying/mmc_mRNAvarying 
        outvec <- c(outvec, cvmc_mRNAvarying)
    }
    
    h5closeAll()
    return(outvec)
}
```

```{r}
#"mRNA_lifeTimes", 1st level: 4839 genes, 2nd: dtype, 3rd: dc/r combo or ctrl
# 4th:50 X 100 ( = 5000 life time data point, some spots will be filled with NAs since not all genes would have so many life times collected), the minimum distribution should have >400 life times collected.
# mml- mean mRNA lifetimes
calculate_gene_sp_mml <- function(geneID, simOutputH5){   
    # outvec has 25 elements. Each element is the cvmc for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes
    currentGene <- geneNames[geneID]

    # loop through the dc/r parameter space
    for(i in 1:nrow(dc.r.df)){  
        ml <- h5read(simOutputH5, paste(currentGene, "mRNA_lifeTimes", dc.r.df$paraCombo[i], sep = "/"))  
        # mRNA life time averaged over simulation time
        mml <- mean(ml, na.rm = TRUE)
        outvec <- c(outvec, mml)
    }
    
    h5closeAll()
    return(outvec)
}
```

```{r}
# vml- variance in mRNA lifetimes
calculate_gene_sp_vml <- function(geneID,simOutputH5){    
    # outvec has 25 elements. Each element is the vml for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes
    currentGene <- geneNames[geneID]

    # loop through the dc/r parameter space
    for(i in 1:nrow(dc.r.df)){  
        ml <- h5read(simOutputH5, paste(currentGene, "mRNA_lifeTimes", dc.r.df$paraCombo[i], sep = "/"))
        # variance in mRNA life time among 5000 values (or maybe less than 5000 filled with NAs)
        vml <- var(as.numeric(ml), na.rm = TRUE)   
        outvec <- c(outvec, vml)
    }
    
    h5closeAll()
    return(outvec)
}
```

```{r}
# cvml- CV in mRNA lifetimes
calculate_gene_sp_cvml <- function(geneID, simOutputH5){    
    # outvec has 25 elements. Each element is the cvml for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes
    currentGene <- geneNames[geneID]

    # loop through the dc/r parameter space
    for(i in 1:nrow(dc.r.df)){  
        # mRNA lifetimes averaged over simulation time
        ml <- h5read(simOutputH5, paste(currentGene, "mRNA_lifeTimes", dc.r.df$paraCombo[i], sep = "/"))  
        mml <- mean(ml, na.rm = TRUE)
        # sdml = standard deviation in mRNA life time among 5000 values (or maybe less than 10000 filled with NAs)
        sdml <- sd(as.numeric(ml), na.rm = TRUE)   
        cvml <- ifelse(mml != 0, sdml/mml, 0) 
        outvec <- c(outvec, cvml)
    }
    
    h5closeAll()
    return(outvec)
}
```

```{r}
# mmdr- mean mRNA decay rates
calculate_gene_sp_mmdr <- function(geneID, simOutputH5){     
    # outvec has 25 elements. Each element is the mmdr for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes
    currentGene <- geneNames[geneID]

    # loop through the dc/r parameter space 
    for(i in 1:nrow(dc.r.df)){  
        # ml = mRNA life time averaged over simulation time
        ml <- h5read(simOutputH5, paste(currentGene, "mRNA_lifeTimes", dc.r.df$paraCombo[i], sep = "/"))  
        # mml = mean mRNA life time averaged over 5000 values (or maybe less than 10000 filled with NAs)
        mml <- mean(ml, na.rm = TRUE)  
        # decay rate = 1/mean lifetime
        mmdr <- 1/mml 
        
        outvec <- c(outvec, mmdr)
    }
    
    h5closeAll()
    return(outvec)
}
```

```{r}
#"gene_sp_mRNAmarkedDecay_perMin", 1st level: 4839 genes, 2nd: dtype, 3rd: dc/r combo or mRNAconstant, 4th: 50 rep X 720 min
# mmmd- mean mRNAs marked for decay 
calculate_gene_sp_mmmd <- function(geneID, simOutputH5){    
    # outvec has 25 elements. Each element is the mmmd (# of mRNA marked for decay averaged over simTimeMin and techReps for certain geneID) for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes
    currentGene <- geneNames[geneID]
    
    # loop through the dc/r parameter space
    for(i in 1:nrow(dc.r.df)){  
        mmd <- h5read(simOutputH5, paste(currentGene, "gene_sp_mRNAmarkedDecay_perMin", dc.r.df$paraCombo[i], sep = "/")) 
        # mean mRNA marked for decay over tech rep and simTimeMin
        mmmd <- mean(mmd)  
        outvec <- c(outvec, mmmd)
    }
    
    h5closeAll()
    return(outvec)
}
```

```{r}
# mmbr =  mean mRNA-averaged bound ribosome
calculate_gene_sp_mmbr <- function(geneID, simOutputH5){    
    # outvec has 25 elements. Each element is the mmbr for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes
    currentGene <- geneNames[geneID]
    
    # loop through the dc/r parameter space
    for(i in 1:nrow(dc.r.df)){  
        ribocount <- h5read(simOutputH5, paste(currentGene, "gene_sp_totBoundRibo_perMin", dc.r.df$paraCombo[i], sep = "/"))
        mRNAcount <- h5read(simOutputH5, paste(currentGene, "gene_sp_mRNAcount_perMin", dc.r.df$paraCombo[i], sep = "/"))
        # total number of bound ribos per gene per minute per rep now divided by real time mRNA count
        mbr <- ribocount/mRNAcount 
        # change all the "Inf"s to "Nan"s
        mbr[!is.finite(mbr)] <- NaN 
        # number of bound ribos per mRNA averaged over time and tech reps
        mmbr <- mean(mbr, na.rm = TRUE) 

        outvec <- c(outvec, mmbr)
    }
    
    h5closeAll()
    return(outvec)
}
```

```{r}
# mtbr =  mean total bound ribosome (total = all mRNAs for a single gene)
calculate_gene_sp_mtbr <- function(geneID, simOutputH5){    
    # outvec has 25 elements. Each element is the mtbr for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes
    currentGene <- geneNames[geneID]
    
    # loop through the dc/r parameter space 
    for(i in 1:nrow(dc.r.df)){  
        tbr <- h5read(simOutputH5, paste(currentGene, "gene_sp_totBoundRibo_perMin", dc.r.df$paraCombo[i], sep = "/"))
        # number of total bound ribos averaged over time and tech reps for each gene
        mtbr <- mean(tbr)
        outvec <- c(outvec, mtbr)
    }
    
    h5closeAll()
    return(outvec)
}
```

```{r}
# mvtbr = mean variance in the total bound ribosomes
calculate_gene_sp_mvtbr <- function(geneID, simOutputH5){    
    # outvec has 25 elements. Each element is the mvtbr for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector() 
    
    # current gene name out of the 4839 genes
    currentGene <- geneNames[geneID]
    
    # loop through the dc/r parameter space
    for(i in 1:nrow(dc.r.df)){   
        vtbr <- h5read(simOutputH5, paste(currentGene, "gene_sp_varianceBoundRibo_perMin", dc.r.df$paraCombo[i], sep = "/"))
        # of bound ribos per mRNA averaged over time and tech reps
        mvtbr <- mean(vtbr) 
        outvec <- c(outvec, mvtbr)
    }
    
    h5closeAll()
    return(outvec)
}
```


```{r}
# mtt = mean translation time (among all ribosomes for a gene)
calculate_gene_sp_mtt <- function(geneID, simOutputH5){  
    # outvec has 25 elements. Each element is the mtt for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector()
    
    # current gene name out of the 4839 genes
    currentGene <- geneNames[geneID]
    
    # loop through the dc/r parameter space
    for(i in 1:nrow(dc.r.df)){
        # att = average translation time among all the ribos
        att <- h5read(simOutputH5, paste(currentGene, "gene_totetimes", dc.r.df$paraCombo[i], sep = "/"))[, 2] 
        mtt <- mean(att)
        outvec <- c(outvec, mtt)
    }
    
    h5closeAll()
    return(outvec)
}
```

```{r}
# mvtt = mean of variance in translation time (var among all ribosomes for a gene)
calculate_gene_sp_mvtt <- function(geneID, simOutputH5){  
    # outvec has 25 elements. Each element is the mvtt for the 1 mRNAconstant + 24 different dc/r combos
    outvec <- vector()
    
    # current gene name out of the 4839 genes
    currentGene <- geneNames[geneID]
    
    # loop through the dc/r parameter space
    for(i in 1:nrow(dc.r.df)){ 
        # vtt = variance in translation time among all the ribos
        vtt <- h5read(simOutputH5, paste(currentGene, "gene_totetimes", dc.r.df$paraCombo[i], sep = "/"))[, 3]
        mvtt <- mean(vtt)
        outvec <- c(outvec, mvtt)
    }
    
    h5closeAll()
    return(outvec)
}
```




#!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!! DON'T RUN FOLLOWING CHUNKS UNLESS THE FEATHER FILES DON'T ALREADY EXIST
### The following chunks utilzie multithreading, make sure the mc.cores are set to approperiate numbers.
```{r}
print(Sys.time())
mpsrList <- mclapply(1:4839, function(x){calculate_gene_sp_mpsr(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# took 2min to run
print(Sys.time())

# ncol = 25, nrow = 4839
mpsrAllgeneAlldcr <- as_tibble(do.call(rbind, mpsrList))

write_feather(mpsrAllgeneAlldcr, path = "../calculatedMetrics/mpsrAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
mvpsList <- mclapply(1:4839, function(x){calculate_gene_sp_mvps(geneID = x, simOutputH5= simOutputH5)}, mc.cores = 45)
# took 2min to run
print(Sys.time())


# ncol = 25, nrow = 4839
mvpsAllgeneAlldcr <- as_tibble(do.call(rbind, mvpsList))

write_feather(mvpsAllgeneAlldcr, path = "../calculatedMetrics/mvpsAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
cvpsList <- mclapply(1:4839, function(x){calculate_gene_sp_cvps(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# 12:44 began 12:49 finish
print(Sys.time())

# ncol = 25, nrow = 4839
cvpsAllgeneAlldcr <- as_tibble(do.call(rbind, cvpsList))

write_feather(cvpsAllgeneAlldcr, path = "../calculatedMetrics/cvpsAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
cpsrList <- mclapply(1:nrow(dc.r.df), function(x){calculate_para_sp_cpsr(pid = x)}, mc.cores = 45)
#parameter id, 1:25, 1 = mRNAconstant, 2:25 = different dc/r combos
# took 2min
print(Sys.time())


# get the cpsr (cellular protein synethesis rates) of 50 tech reps for all the parameter combos
cpsrAlldcrAllrep <- as_tibble(t(data.frame(do.call(rbind, cpsrList))))  # ncol=25 pid, nrow=50 tech reps
colnames(cpsrAlldcrAllrep) <- dc.r.df$paraCombo

write_feather(cpsrAlldcrAllrep, path = "../calculatedMetrics/cpsrAlldcrAllrep.feather")

```

```{r}
print(Sys.time())
# pid = parameter id = 1:25, 1 = mRNAconstant, 2:25 = different dc/r combos, output is a number
mcpsrList <- mclapply(1:nrow(dc.r.df), function(x){calculate_para_sp_mcpsr(pid = x)}, mc.cores = 45)
# took 2min
print(Sys.time())

# mcpsrAlldcr is a vector of 25 numbers
mcpsrAlldcr <- as_tibble(data.frame(MCPSR = unlist(mcpsrList),
                       paraID = dc.r.df$paraID))

write_feather(mcpsrAlldcr, path = "../calculatedMetrics/mcpsrAlldcr.feather")
```

```{r}
print(Sys.time())
mteList <- mclapply(1:4839, function(x){calculate_gene_sp_mte(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
print(Sys.time())

# ncol = 25, nrow = 4839
mteAllgeneAlldcr <- as_tibble(do.call(rbind, mteList))

write_feather(mteAllgeneAlldcr, path = "../calculatedMetrics/mteAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
mvteList <- mclapply(1:4839, function(x){calculate_gene_sp_mvte(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# took 3 min
print(Sys.time())

# ncol = 25, nrow = 4839
mvteAllgeneAlldcr <- as_tibble(do.call(rbind, mvteList))

write_feather(mvteAllgeneAlldcr, path = "../calculatedMetrics/mvteAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
mmcList <- mclapply(1:4839, function(x){calculate_gene_sp_mmc(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# took 2min to run
print(Sys.time())

# ncol = 25, nrow = 4839
mmcAllgeneAlldcr <- as_tibble(do.call(rbind, mmcList))

write_feather(mmcAllgeneAlldcr, path = "../calculatedMetrics/mmcAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
mvmcList <- mclapply(1:4839, function(x){calculate_gene_sp_mvmc(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# took 2min to run
print(Sys.time())

# ncol = 25, nrow = 4839
mvmcAllgeneAlldcr <- as_tibble(do.call(rbind, mvmcList))

write_feather(mvmcAllgeneAlldcr, path = "../calculatedMetrics/mvmcAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
cvmcList <- mclapply(1:4839, function(x){calculate_gene_sp_cvmc(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# 12:44 began 12:49 finish
print(Sys.time())

# ncol = 25, nrow = 4839
cvmcAllgeneAlldcr <- as_tibble(do.call(rbind, cvmcList))

write_feather(cvmcAllgeneAlldcr, path = "../calculatedMetrics/cvmcAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
mmlList <- mclapply(1:4839, function(x){calculate_gene_sp_mml(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# took 2min to run
print(Sys.time())

# ncol = 25, nrow = 4839
mmlAllgeneAlldcr <- as_tibble(do.call(rbind, mmlList))

write_feather(mmlAllgeneAlldcr, path = "../calculatedMetrics/mmlAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
vmlList <- mclapply(1:4839, function(x){calculate_gene_sp_vml(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# took 2min to run
print(Sys.time())

# ncol = 25, nrow = 4839
vmlAllgeneAlldcr <- as_tibble(do.call(rbind, vmlList))

write_feather(vmlAllgeneAlldcr, path = "../calculatedMetrics/vmlAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
cvmlList <- mclapply(1:4839, function(x){calculate_gene_sp_cvml(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# took 2min to run
print(Sys.time())

# ncol = 25, nrow = 4839
cvmlAllgeneAlldcr <- as_tibble(do.call(rbind, cvmlList))

write_feather(cvmlAllgeneAlldcr, path = "../calculatedMetrics/cvmlAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
mmdrList <- mclapply(1:4839, function(x){calculate_gene_sp_mmdr(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# took 2min to run
print(Sys.time())

# ncol = 25, nrow = 4839
mmdrAllgeneAlldcr <- as_tibble(do.call(rbind, mmdrList))

write_feather(mmdrAllgeneAlldcr, path = "../calculatedMetrics/mmdrAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
mmmdList <- mclapply(1:4839, function(x){calculate_gene_sp_mmmd(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# took 3 min
print(Sys.time())

# ncol = 25, nrow = 4839
mmmdAllgeneAlldcr <- as_tibble(do.call(rbind, mmmdList))

write_feather(mmmdAllgeneAlldcr, path = "../calculatedMetrics/mmmdAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
mmbrList <- mclapply(1:4839, function(x){calculate_gene_sp_mmbr(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# took 2min to run
print(Sys.time())

# ncol = 25, nrow = 4839
mmbrAllgeneAlldcr <- as_tibble(do.call(rbind, mmbrList))

write_feather(mmbrAllgeneAlldcr, path = "../calculatedMetrics/mmbrAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
mtbrList <- mclapply(1:4839, function(x){calculate_gene_sp_mtbr(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# took 2min to run
print(Sys.time())

# ncol = 25, nrow = 4839
mtbrAllgeneAlldcr <- as_tibble(do.call(rbind, mtbrList))

write_feather(mtbrAllgeneAlldcr, path = "../calculatedMetrics/mtbrAllgeneAlldcr.feather")
```

```{r}
print(Sys.time())
mvtbrList <- mclapply(1:4839, function(x){calculate_gene_sp_mvtbr(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# took 2min to run
print(Sys.time())

# ncol = 25, nrow = 4839
mvtbrAllgeneAlldcr <- as_tibble(do.call(rbind, mvtbrList))

write_feather(mvtbrAllgeneAlldcr, path = "../calculatedMetrics/mvtbrAllgeneAlldcr.feather")
```


```{r}
print(Sys.time())
mttList <- mclapply(1:4839, function(x){calculate_gene_sp_mtt(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# took 2min to run
print(Sys.time())

# ncol = 25, nrow = 4839
mttAllgeneAlldcr <- as_tibble(do.call(rbind, mttList))

write_feather(mttAllgeneAlldcr, path = "../calculatedMetrics/mttAllgeneAlldcr.feather")

```


```{r}
print(Sys.time())
mvttList <- mclapply(1:4839, function(x){calculate_gene_sp_mvtt(geneID = x, simOutputH5 = simOutputH5)}, mc.cores = 45)
# took 2min to run
print(Sys.time())

# ncol = 25, nrow = 4839
mvttAllgeneAlldcr <- as_tibble(do.call(rbind, mvttList))

write_feather(mvttAllgeneAlldcr, path = "../calculatedMetrics/mvttAllgeneAlldcr.feather")

```

# calculating the mean number of free ribos (averaged over simTimeMin and techReps) for all dc.r.combos
```{r}
# MFR = mean free ribo
# RMFR = relative mean free ribo
mfr <- c() 
# loop through the dc/r parameter space
for(i in 1:nrow(dc.r.df)){ 
    # free_ribo averaged by simTimeMin and techReps
    mfr <- c(mfr, mean(h5read(simOutputH5, paste("free_ribo_tRNA_perMin", dc.r.df$paraCombo[i], "ribo", sep = "/")), na.rm = TRUE))
}

rmfr <- mfr/mfr[1]
rmfr_df <- as_tibble(data.frame(RMFR = rmfr,
                      MFR = mfr,
                      paraID = dc.r.df$paraID))

```


# getting all input features and output results together in allSimInputOutput
```{r}
# convert any input dataframes (e.g. mteAllgeneAlldcr) to long format
df.convert.fun <- function(x){   
    inputdf = calculatedMetrics.list[[x]]
    dfname = str_sub(toupper(names(calculatedMetrics.list)[[x]]), 1, -14)

    inputdf %>%
        mutate(ORF = simInputFeatures$ORF) %>%
        gather_(key_col = "paraID", value_col = dfname, gather_cols = colnames(inputdf)[1:nrow(dc.r.df)])

}

# Read in the files that conatins "AllgeneAlldcr.feather", these files can be left_joined by "paraID" and "ORF"
files.locs <- dir("../calculatedMetrics", pattern = "AllgeneAlldcr.feather", full.names = TRUE)

# read_feather for all the files in this location
calculatedMetrics.list <- sapply(files.locs, read_feather, simplify = FALSE)  
# exclude the path and only extract the data frame names, e.g. "mmcAllgeneAlldcr"
names(calculatedMetrics.list) <- str_sub(names(calculatedMetrics.list), nchar("../calculatedMetrics/")+1, -9)   


# divide mte of 25 dc/r combo by mte of the mRNAconstant = rmte, relative mean translation efficiency
# X1 = mRNA constant, X2:X25 = mRNAvarying
# dim(rmteAllgeneAlldcr) = 4839 X 25
calculatedMetrics.list$rmpsrAllgeneAlldcr <- as_tibble(t(apply(calculatedMetrics.list$mpsrAllgeneAlldcr, 1, function(x){x[1:nrow(dc.r.df)]/x[1]})))
calculatedMetrics.list$rmvpsAllgeneAlldcr <- as_tibble(t(apply(calculatedMetrics.list$mvpsAllgeneAlldcr, 1, function(x){x[1:nrow(dc.r.df)]/x[1]})))
calculatedMetrics.list$rcvpsAllgeneAlldcr <- as_tibble(t(apply(calculatedMetrics.list$cvpsAllgeneAlldcr, 1, function(x){x[1:nrow(dc.r.df)]/x[1]}))) #x[1] is for mRNAconstant
calculatedMetrics.list$rmteAllgeneAlldcr <-  as_tibble(t(apply(calculatedMetrics.list$mteAllgeneAlldcr, 1, function(x){x[1:nrow(dc.r.df)]/x[1]}))) #x[1] is for mRNAconstant
calculatedMetrics.list$rmvteAllgeneAlldcr <- as_tibble(t(apply(calculatedMetrics.list$mvteAllgeneAlldcr, 1, function(x){x[1:nrow(dc.r.df)]/x[1]})))
calculatedMetrics.list$cvmcAllgeneAlldcr <- sqrt(calculatedMetrics.list$mvmcAllgeneAlldcr)/calculatedMetrics.list$mmcAllgeneAlldcr  # mRNA count coefficient of variation
calculatedMetrics.list$rmmcAllgeneAlldcr <- as_tibble(t(apply(calculatedMetrics.list$mmcAllgeneAlldcr, 1, function(x){x[1:nrow(dc.r.df)]/x[1]})))
calculatedMetrics.list$rmmdrAllgeneAlldcr <- as_tibble(do.call(rbind, lapply(1:4839, function(geneID){calculatedMetrics.list$mmdrAllgeneAlldcr[geneID, ]/simInputFeatures$mRNADecRateNeymotin_sec[geneID]})))  # for X1 this will be 0,  for the ~30 genes whose GreshamDecRate = 0 this will be NaN b/c 0/0 = NaN
calculatedMetrics.list$rmttAllgeneAlldcr <- as_tibble(t(apply(calculatedMetrics.list$mttAllgeneAlldcr, 1, function(x){x[1:nrow(dc.r.df)]/x[1]})))
calculatedMetrics.list$rmvttAllgeneAlldcr <- as_tibble(t(apply(calculatedMetrics.list$mvttAllgeneAlldcr, 1, function(x){x[1:nrow(dc.r.df)]/x[1]})))
calculatedMetrics.list$mhlAllgeneAlldcr <- as_tibble(log(2)/calculatedMetrics.list$mmdrAllgeneAlldcr) # mean half life
calculatedMetrics.list$mmdpAllgeneAlldcr <- as_tibble(calculatedMetrics.list$mmmdAllgeneAlldcr/calculatedMetrics.list$mmcAllgeneAlldcr)
calculatedMetrics.list$mrdAllgeneAlldcr <- as_tibble(calculatedMetrics.list$mmbrAllgeneAlldcr/simInputFeatures$geneLength_codon) #ribosome density=number of ribosomes N divided by the CDS length L
calculatedMetrics.list$rmrdAllgeneAlldcr <- as_tibble(t(apply(calculatedMetrics.list$mrdAllgeneAlldcr, 1, function(x){x[1:nrow(dc.r.df)]/x[1]}))) #x[1] is for mRNAconstant
calculatedMetrics.list$rmtbrAllgeneAlldcr <- as_tibble(t(apply(calculatedMetrics.list$mtbrAllgeneAlldcr, 1, function(x){x[1:nrow(dc.r.df)]/x[1]}))) #x[1] is for mRNAconstant
calculatedMetrics.list$rmvtbrAllgeneAlldcr <- as_tibble(t(apply(calculatedMetrics.list$mvtbrAllgeneAlldcr, 1, function(x){x[1:nrow(dc.r.df)]/x[1]}))) #x[1] is for mRNAconstant

# convert all the calculatedMetrics.list dataframes to long format
calculatedMetricsLong.list <- lapply(1:length(calculatedMetrics.list), df.convert.fun)

mcpsrAlldcr <- read_feather("../calculatedMetrics/mcpsrAlldcr.feather")

# Join all the dataframes in calculatedMetrics.list
allSimInputOutput <- 
    calculatedMetricsLong.list %>%
        purrr::reduce(left_join, by = c("paraID" = "paraID", "ORF" = "ORF")) %>%
        left_join(., mcpsrAlldcr, "paraID") %>%
        left_join(., rmfr_df, by = "paraID") %>%
        left_join(., dc.r.df, "paraID") %>%
        left_join(., simInputFeatures, by = "ORF")

write_feather(allSimInputOutput, path = "../allSimInputOutput.feather")

```

```{r}
cpsrAlldcr
mcpsrAlldcr
```


```{r}

allSimInputOutput %>%
    filter(ORF == "YBR208C")

allSimInputOutput %>%
    filter(ORF == "YJR109C")

allSimInputOutput %>%
    filter(ORF == "YAL001C")

allSimInputOutput %>%
    filter(ORF == "YAL034W-A")
```