---
title: "Figure and captions"
output: pdf_document
geometry: "left=1cm,right=1cm,top=1.5cm,bottom=1.5cm"
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r echo = FALSE}
# Prevent printing of warnings and such in the HTML
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center", echo = FALSE, dev = "cairo_pdf")
```

```{r}
library(tidyverse)
library(patchwork)

# these are the locations of the figure rds, they're read in with no storage to avoid placing a large object in RAM
figRdsLocs <- dir("../figures", full.names = TRUE, pattern = ".rds")

names(figRdsLocs) <- stringr::str_remove(figRdsLocs, "../figures/") %>% 
  str_remove(".rds")
```

### Model validation plot
```{r fig.height=4, fig.width=6}
read_rds(figRdsLocs[names(figRdsLocs) == "validationPlot"])
```
Validation plot: \newpage

### FigureS0
```{r fig.height=4, fig.width=6}
read_rds(figRdsLocs[names(figRdsLocs) == "figS0_a"])
```
figS0_a: 

```{r fig.height=4, fig.width=6}
read_rds(figRdsLocs[names(figRdsLocs) == "figS0_b"])
```
figS0_b: \newpage


```{r fig.height=4, fig.width=6}
read_rds(figRdsLocs[names(figRdsLocs) == "figS0_c"])
```
figS0_c: 

```{r fig.height=4, fig.width=6}
read_rds(figRdsLocs[names(figRdsLocs) == "figS0_d"])
```
figS0_d: \newpage


```{r fig.height=4, fig.width=6}
read_rds(figRdsLocs[names(figRdsLocs) == "figS0_e"])
```
figS0_e: 

```{r fig.height=4, fig.width=6}
read_rds(figRdsLocs[names(figRdsLocs) == "figS0_f"])
```
figS0_f: \newpage


```{r fig.height=4, fig.width=6}
read_rds(figRdsLocs[names(figRdsLocs) == "figS0_g"])
```
figS0_g: 

```{r fig.height=4, fig.width=6}
read_rds(figRdsLocs[names(figRdsLocs) == "figS0_h"])
```
figS0_h: \newpage


```{r fig.height=4, fig.width=6}
read_rds(figRdsLocs[names(figRdsLocs) == "figS0_i"])
```
figS0_i: 

```{r fig.height=4, fig.width=6}
read_rds(figRdsLocs[names(figRdsLocs) == "figS0_j"])
```
figS0_j: \newpage



### FigureS1
```{r fig.height=6, fig.width=10}
read_rds(figRdsLocs[names(figRdsLocs) == "figureS1"])
```
FigureS1: \newpage


### FigureS2
```{r fig.height=6, fig.width=10}
read_rds(figRdsLocs[names(figRdsLocs) == "figureS2"])
```
FigureS2: \newpage


### FigureS3
```{r fig.height = 3, fig.width = 8}
read_rds(figRdsLocs[names(figRdsLocs) == "figureS3"])
```
FigureS3: 

### FigureS4
```{r fig.height = 4, fig.width = 6}
read_rds(figRdsLocs[names(figRdsLocs) == "figureS4"])
```
FigureS4: \newpage


### FigureS5
```{r fig.height = 4, fig.width = 6}
read_rds(figRdsLocs[names(figRdsLocs) == "figureS5"])
```
FigureS5: 

### FigureS6
```{r fig.height = 3, fig.width = 8}
read_rds(figRdsLocs[names(figRdsLocs) == "figureS6"])
```
FigureS6: \newpage


### FigureS7
```{r fig.height = 4, fig.width = 7}
read_rds(figRdsLocs[names(figRdsLocs) == "figureS7"])
```
FigureS7: 

### FigureS8
```{r fig.height = 4, fig.width = 7}
read_rds(figRdsLocs[names(figRdsLocs) == "figureS8"])
```
FigureS8: \newpage

### FigureS9
```{r fig.height = 7, fig.width = 10}
read_rds(figRdsLocs[names(figRdsLocs) == "figureS9"])
```
FigureS9: \newpage

### FigureS10
```{r fig.height = 7, fig.width = 10}
read_rds(figRdsLocs[names(figRdsLocs) == "figureS10"])
```
FigureS10: \newpage
