---
title: "R Notebook"
output: html_notebook
---

Load libraries and public variables
```{r, include = FALSE} 
# This chunk will be evaluated but the output is suppressed
rmarkdown::render("../../publicVariables/createPublicVariables.Rmd")
```

```{r}
# File that contains simulation output data in a HDF5 directory structure specified in "/publicVariables/createPublicVariables.Rmd"
simOutputH5 <- paste0("../../dataProcessing/simOutputH5Data/", experimentID, "_output.h5")

# A comprehensive table that contains all simulation input and calculated output metrics
allSimInputOutput <- read_feather("../allSimInputOutput.feather")

# This folder contains raw output files from the simulations. This is needed for calculate_para_sp_mcpsr()
rawSimOutFolder <- "/team/batch_SMOTNT/experiment1_output"

# Where figures are stored
figureBaseFolder <- "../figures/"
```

```{r}
# protein synthesized per minute
PS_byTimeRep <- function(gName1, gName2, gName3)
{
    protSynRate_ls <- lapply(c(gName1, gName2, gName3), function(x){
      tibble(rbind(data.frame(t(h5read(simOutputH5, paste0(gsub("-", "", x), "/gene_sp_countElng_perMin/dc_1_r_0")))), data.frame(t(h5read(simOutputH5, paste0(gsub("-", "", x), "/gene_sp_countElng_perMin/mRNAconstant"))))))})
    #each tibble is 1440 X 50 reps, 1440 = 720min for each of dc_1_r_0 or ctrl
    
    coeff <- lapply(c(gName1, gName2, gName3), function(x){((allSimInputOutput%>%filter(ORF==x, paraCombo=="dc_1_r_0"))$MPSR)/((allSimInputOutput%>%filter(ORF==x, paraCombo=="dc_1_r_0"))$mRNAabundance)}) # coeff in order to transform the sec_axis
    
    # The protein synthesis rates are transformed, in order to transform the sec_axis below without a problem (too complicated to explain here)..The main obstacle is that by using facet_wrap, it is impossible to customize the second y axis for each facet, without transforming the values beforehand like below.
    protSynRate_ls[[1]] <- protSynRate_ls[[1]]*coeff[[2]]*coeff[[3]]
    protSynRate_ls[[2]] <- protSynRate_ls[[2]]*coeff[[1]]*coeff[[3]]
    protSynRate_ls[[3]] <- protSynRate_ls[[3]]*coeff[[1]]*coeff[[2]]
    
    protSynRepTime <- do.call(rbind, protSynRate_ls)    # 4320 X 50, 4320=1440x3
  
    colnames(protSynRepTime) <- paste0("rep", 1:numTechReps)
    
    
    mRNAabundRealTime <- lapply(c(gName1, gName2, gName3), function(x){c((h5read(simOutputH5, paste0(gsub("-", "", x), "/gene_sp_mRNAcount_perMin/dc_1_r_0")))[1, ],  rep(simInputFeatures$mRNAabundance[which(simInputFeatures$ORF==x)], simTimeMin))})
    mRNAabundRealTime <- do.call(c, mRNAabundRealTime)
    
    p1 <- protSynRepTime %>%
      mutate(dc.r.paraCombo = rep(c(rep("dc_1_r_0", simTimeMin), rep("mRNAconstant", simTimeMin)), 3), 
             time = rep(c(1:simTimeMin, 1:simTimeMin), 3), 
             gName = c(rep(gName1, simTimeMin*2), rep(gName2, simTimeMin*2), rep(gName3, simTimeMin*2)), 
             mRNArealTime = mRNAabundRealTime) %>%
      pivot_longer(names_to = "reps",values_to = "PS", cols = 1:numTechReps) %>%
      ggplot(aes(x=time,  color=dc.r.paraCombo)) +
      geom_line(aes(y = mRNArealTime)) +
      geom_line(aes(y = PS/(coeff[[1]]*coeff[[2]]*coeff[[3]])), alpha=0.5) +
      facet_wrap(~factor(gName, levels=c(gName1, gName2, gName3)), scale="free", nrow=3) +
      theme_bw() +
      theme(text = element_text(size = 13),
            legend.position = "top",
            legend.justification = "left",
            strip.background = element_blank(),
            panel.grid = element_blank(),
            plot.margin = margin(r = 0, unit = "pt")
            ) +
      labs(x="Time(min)",
           y="Protein synthesized within a minute") +
      scale_color_manual(name="", labels=c("mRNAconstant"="mRNAconstant", "dc_1_r_0"="100% CMD"), values = c("#1F78B4", "#FF7F00")) +
      scale_y_continuous(expand = c(0, 0), name="Real-time mRNA abundance (single replicate)", sec.axis = sec_axis(~.*(coeff[[1]]*coeff[[2]]*coeff[[3]]), name = "Protein synthesized per minute (50 replicates)"))  # remove the extra space on the axis

    
    p2 <- protSynRepTime %>%
      mutate(dc.r.paraCombo = rep(c(rep("dc_1_r_0", simTimeMin), rep("mRNAconstant", simTimeMin)), 3),
             time = rep(c(1:simTimeMin, 1:simTimeMin), 3),
             gName = c(rep(gName1, simTimeMin*2), rep(gName2, simTimeMin*2), rep(gName3, simTimeMin*2))) %>%
      pivot_longer(names_to = "reps", values_to = "PS", cols = 1:numTechReps) %>%
      ggplot(aes(x=PS, fill=dc.r.paraCombo, color = dc.r.paraCombo)) +
      geom_density(alpha=0.5) +
      facet_wrap(~factor(gName, levels=c(gName1, gName2, gName3)), scale="free", nrow=3) +
      theme_bw() +
      theme(text = element_text(size = 13),
            legend.position = "none",
            axis.ticks.x = element_blank(),
            axis.ticks.y = element_blank(),
            axis.text.x = element_blank(),
            axis.text.y = element_blank(),
            strip.text = element_text(color="white"),  # can't use element_blank() here because it gives error when p1+p2
            strip.background = element_blank(),
            panel.border = element_blank(),
            panel.background = element_blank(),
            panel.grid = element_blank(),
            plot.margin = margin(r = 0, unit = "pt")
    ) +
      labs(x="") +
      scale_fill_manual(name="", labels=c("mRNAconstant"="mRNAconstant", "dc_1_r_0"="Co-trans 100%"), values = c("#1F78B4", "#FF7F00")) +
      scale_color_manual(values = c("#1F78B4", "#FF7F00")) +
      scale_y_continuous(expand = c(0, 0)) +
      coord_flip()


    p1 + p2 + plot_layout(ncol = 2, width = c(4, 0.6))
}

```

```{r}
# These three genes each has their mRNAconstant <, =, > mRNAvarying    mRNAabundance high
figure1a <- PS_byTimeRep("YOR369C", "YHR064C", "YKL182W") 

figure1c <- allSimInputOutput %>%
  filter(paraCombo=="mRNAconstant"|paraCombo=="dc_1_r_0", ORF!="YER053C-A") %>%
  ggplot(aes(x=MPSR, y=MVPS, color = paraCombo)) +
  geom_point(size = 0.5) +
  geom_abline(linetype="dashed") +
  theme_bw() +
  theme(text = element_text(size = 13),
            legend.position = "none",
            strip.text = element_text(color="white"),  # can't use element_blank() here because it gives error when p1+p2
            plot.margin = margin(r = 0, unit = "pt"),
            panel.grid = element_blank()
    ) +
      labs(x = 'Mean of protein synthesis rate'~(min^-1),
           y = 'Variance of protein synthesis rate'~(min^-1)) +
      scale_color_manual(name="", labels=c("mRNAconstant"="mRNAconstant", "dc_1_r_0"="Co-trans 100% RiboProtect 0"), values = c("#1F78B4", "#FF7F00")) +
      geom_smooth(method="lm") +
      stat_cor(aes(label=paste(..r.label..,  sep = "~`,`~")), label.x.npc = 0.7, label.y.npc = 0.4, size=5) +
      scale_x_log10() +
      scale_y_log10()

#################################################################
figure1d <- allSimInputOutput %>%
  filter(paraCombo=="dc_1_r_0", ORF!="YER053C-A") %>%
  ggplot(aes(x=RMPSR, y=RMVPS, color = paraCombo)) +
  geom_point(size = 0.5) +
  theme_bw() +
  theme(text = element_text(size = 13),
            legend.position = "none",
            strip.text = element_text(color="white"),  # can't use element_blank() here because it gives error when p1+p2
            plot.margin = margin(r = 0, unit = "pt"),
            panel.grid = element_blank()
    ) +
      labs(x ="Rel. mean of protein synthesis rate",
           y = "Rel. variance of protein synthesis rate") +
      stat_cor(aes(label = paste(..r.label..,  sep = "~`, `~")), label.x.npc = 0.2, label.y.npc = 0.5, size=5) +
      scale_x_log10() + # not able to add the inset if i keep the log scales for either x or y
      scale_y_log10() +
      scale_color_manual(name="", labels=c("Co-trans 100% RiboProtect 0"), values = c("black"))

#################################################################
figure1e <- allSimInputOutput %>%
  filter(paraCombo=="dc_1_r_0", ORF!="YER053C-A") %>% #remove the gene with the 450min HL
  mutate(intCtrlOrNot = ifelse(mRNADecRateNeymotin_sec==0,"intCtrl","nonCtrl")) %>% 
  select("ORF", "RMPSR") %>% 
  ggplot() +
  geom_density(aes(x=RMPSR, color = intCtrlOrNot)) +
  theme_bw() +
  theme(text = element_text(size = 13),
        legend.position = c(0.6, 0.6),
        legend.background = element_blank(),
        axis.ticks.y = element_blank(),
        plot.margin = margin(r = 0, unit = "pt"),
        panel.grid = element_blank()) +
    labs(x = "Rel. mean protein synthesis rate") +
    scale_color_manual(values = c("RMVPS"="black")) +  
    scale_x_log10() +
    geom_vline(xintercept = 1, linetype = "dashed") 

#################################################################
figure1f <- allSimInputOutput %>%
  filter(paraCombo=="dc_1_r_0", ORF!="YER053C-A") %>% #remove the gene with the 450min HL
  select("ORF", "RMVPS") %>% 
  ggplot() +
  geom_density(aes(x=RMVPS)) +
  theme_bw() +
  theme(text = element_text(size = 13),
        legend.position = c(0.6, 0.6),
        legend.background = element_blank(),
        axis.ticks.y = element_blank(),
        plot.margin = margin(r = 0, unit = "pt"),
        panel.grid = element_blank()) +
    labs(x = "Rel.variance in protein synthesis rates") +
    #scale_color_manual(values = c("RMVPS"="black")) +  
    scale_color_manual(values = c("intCtrl"="blue", "nonCtrl"="black")) +  
    scale_x_log10() +
    geom_vline(xintercept = 3, linetype = "dashed") 

```


Relative mean of protein synthesis rate = for each gene, the ratio between mean of protein synthesis rate for dc_1_r_0 and mRNAconstant (mean among 720min X 50rep numbers).
Relative variance of protein synthesis rate = for each gene, the ratio between variance of protein synthesis rate for dc_1_r_0 and mRNAconstant (variance among 720min X 50rep numbers).
```{r fig.height = 10, fig.width=15}
################################################################ Fig. 1
figure1 <- figure1a | ((figure1c + figure1d)/(figure1e + figure1f))
figure1 + plot_annotation(
  title = "Figure 1",
  tag_levels = 'A')

ggsave(paste0(figureBaseFolder,"a_fig_1.png"), dpi=200, height = 10, width = 15)


```

