---
title: "R Notebook"
output: html_notebook
---

A dispersion test to check if the mRNA count in Fig 1A follows a Poisson distribution. It relies on the fact that the Poisson distribution's mean is equal to its variance, and the the ratio of the variance to the mean in a sample of n counts from a Poisson distribution should follow a Chi-square distribution with n-1 degrees of freedom.

This allows you to reject the null hypothesis that your data are Poisson distributed if the p value is <0.05. If the p value is above 0.05, you could accept that the data followed a Poisson distribution.
```{r}
dispersion_test <- function(x) 
{
  res <- 1-2 * abs((1 - pchisq((sum((x - mean(x))^2)/mean(x)), length(x) - 1))-0.5)

  cat("Dispersion test of count data:\n",
      length(x), " data points.\n",
      "Mean: ",mean(x),"\n",
      "Variance: ",var(x),"\n",
      "Probability of being drawn from Poisson distribution: ", 
      round(res, 3),"\n", sep = "")

  invisible(res)
}
```

```{r}
x <- h5read(simOutputH5, "YOR369C/gene_sp_mRNAcount_perMin/dc_1_r_0")[1,]
dispersion_test(x)
```

